---
title: Time Series Lessons
author: Joshua Lammon
date: '2017-10-08'
slug: time-series-lessons
categories:
  - Data Science
tags:
  - Data
  - MLB
  - Fun
---
Welcome back to Data Scientology. Today we will be looking at how to create a time series plot on some Babe Ruth's HR career homerun totals.

First, like always, we must upload necessary libraries needed to perform the time series function. The three libraries we need are as follows:
```{r warning=FALSE, message=FALSE}
library(Lahman)
library(sqldf)
library(ggplot2)
```

Next, we must execute a query to extract the data we want to create a time-series plot on. We want each year Babe Ruth played in the MLB and we also want his homerun totals. Luckily, they are both located in the Batting table so we just want to pull those two fields from the playerID "ruthba01." 
```{r}
query<-"SELECT yearID,HR
FROM Batting
WHERE playerID='ruthba01'"
result<-sqldf(query)
```

Next, we want to write a "ggplot" function that will allow us to create the time series plot and visualize the data. To manipulate the graph, we must use the "aes" function which allows us to change the graph's aesthetics. We can choose many different colors to add personality to our graphs, and we can select different sizes for the lines and individual points in our time series graph. To manually edit the title, we have to add a line indicated by the "+" symbol and the use the "ggtitle" function. We can also select the x and y-axis titles by changing the "xlab" and "ylab" functions.
```{r}
#visualize the data
ggplot()+
  geom_line(data=result,aes(x=yearID, y=HR), size=1.1, color="red")+
  geom_point(data=result,aes(x=yearID,y=HR), size=2)+
  ggtitle("Ruth's Homerun Totals Through the Years")+
  xlab("Year")+
  ylab("Homeruns")
```

As you can see in the visualization above, we have illustrated Babe Ruth's annual homerun totals over his career. Babe Ruth is an MLB legend and is considered one of the pioneers of the sport. 

Hopefully this blog post was helpful and you now understand how to create a simple time series plot using R Markdown and ggplot. Please continue to check back in for future posts relating to data science. Feel free to contact me if you have any questions or comments. 
